### 2023-11-13 outputs/paper/paper.qmd
Change figures and add discussion part

---
title: "Public Security Concern: Bicycle Theft in Toronto Shows a Seasonal and Geographical Pattern"
author: "Yufei Liu"
thanks: "Code and data are available at: https://github.com/Florence-Liu/bicycle-theft"
date: today
date-format: long
abstract: "Bicycle theft has been a serious issue for city safety in Toronto. The analysis uses data on opendatatoronto portal including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service. We find a decreasing trend and low recovery rates in overall offences in recent years even people tend to report immediately, a seasonal pattern that more offences occurred in summer, and a geographical pattern that outdoor bicycle thefts decreased. It implies possible ineffectiveness of tracking and security management, and further work could look at influencing factors contributing to the patterns."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Cycling is one of the most popular transportation methods and relaxing activities in Toronto and contributes to reducing traffic congestion. The City of Toronto supports cycling by building infrastructures, such as bike lanes and bike parking facilities, and create a cycling-friendly environment [@citeThecity]. However, together with the prevalence of cycling, bicycle theft became a major concern for cyclists and the city safety in Toronto.

In this paper, we use R [@citeR] to analyze `bicycle thefts` dataset [@citeData] in `opendatatoronto` package [@citeOpendatatoronto] to discover any potential time and geographical patterns in bicycle theft occurrence in Toronto from 2014 to 2022. We explore time and location variables as well as recovery rate and report habit through bar plots, scatter plots, and tables. We find that the overall occurrence of bicycle theft in Toronto decreased in recent years with a potential seasonal pattern that more thefts occurred in summer, and a geographical pattern that outdoor thefts decreased, along with relatively law recovery rates and immediate report habit. Biases exist in the dataset and may influence some analysis. Further work could explore possible reasons behind the patterns combining with other dataset.

The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions; @sec-discussion discusses findings from data visualization and implications.

# Data {#sec-data}

## Data introduction  {#sec-data_introduction}

The dataset used in this analysis was obtained from Toronto's Open Data Portal and can be accessed from the package `opendatatoronto` [@citeOpendatatoronto]. The dataset `Bicycle Thefts` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022 [@citeData]. Since the occurrence of bicycle theft was self-reported, there may exist reporting bias that the reporter may choose to hide or inaccurately report some information [@citeTorontopoliceservice]. Also, the location information is blurred due to privacy consideration.  We use the latest `csv` version refreshed on Apr 4, 2023 in this analysis.

We will use `R` to do the analysis in this report [@citeR]. After acquiring the dataset from package `opendatatoronto` [@citeOpendatatoronto], we used package `tidyverse` [@citeTidyverse] and package `janitor` [@citeJanitor] to clean the data and select variables for further analysis.

We selected variables from the raw data [@citeData]:

* `occ_year`: Year offence occurred

* `occ_month`: Month offence occurred

* `occ_date`: Date offence occurred

* `report_date`: Date offence reported

* `premises_type`: Premises type of offence

* `status`: Whether the bicycle is recovered, stolen, or unknown

To analyze the relationship between occurrence date and report date of offence, we constructed a new numerical variable `difference` to represent the difference in days between the report date and occurrence date, and filter any offences whose difference is larger than 365, that means we will not analyze offence that is not reported within one year from the occurrence date. The number of offence that reports long time later is quite small and may not be representative due to possible inaccurate report information.

After changing the type and name of some variables, filtering offence occurred between 2014 and 2022, and removing missing values, we selected our final variable for analysis:

* `occ_year`: Year offence occurred

* `occ_month`: Month offence occurred

* `difference`: Difference in days between occurrence date and report date

* `location`: Premises type of offence

* `bicycle_status`: Whether the bicycle is recovered, stolen, or unknown

## Data visualization {#sec-data_visualization}

In the analysis, we will use package `knitr` [@citeKnitr], package `here` [@citeHere], and package `formattable` [@citeFormattable] to load the cleaned data and make figures and tables to explore relationship between variables.

```{r}
#| echo: false
#| message: false
#| warning: false

#### Workspace setup ####
library(tidyverse)
library(knitr)
library(formattable)

#### Read in the cleaned data ####
data <- read_csv(here::here("outputs/data/cleaned_bicycle_theft.csv"), 
                 show_col_types = FALSE,
                 col_types = cols(occ_year = col_character(),
                                  occ_month = col_character()))
```

```{r}
#| label: fig-yearly_count
#| fig-cap: Number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
#| echo: false
#| warning: false
#| message: false

data |> 
  mutate(occ_month = factor(occ_month, levels = month.name)) |> 
  ggplot(aes(x=as.factor(occ_year))) +
  geom_bar(color="darkgreen", fill="pink2") +
  labs(x = "Occurrence Year", y = "Number of occurrence", fill = "Occurrence Month") +
  theme_minimal()
```

@fig-yearly_count shows the total number of bicycle thefts in Toronto from 2014 to 2022. We could find that the total number of offence in a year increased before 2018 and began decrease since 2018. The change is relatively not significant within 1000, and the number of offence in 2022 is similar to that in 2014.

@fig-yearly_count_by_month illustrates potential seasonal pattern that there are much fewer bicycle thefts in winter, eg. December, January, and February. The offence increases a lot in other seasons especially in summer, eg. June, July, and August. It also shows that the season in which bicycles were stolen differs from year to year, but overall centers in the summer seasons and the seasonal difference is significant.

```{r}
#| label: fig-yearly_count_by_month
#| fig-cap: Monthly number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
#| echo: false
#| warning: false
#| message: false

data |> 
  mutate(occ_month = factor(occ_month, levels = month.name)) |> 
  group_by(occ_year, occ_month) |>
  summarise(n=n()) |>
  ggplot(aes(x = occ_month, y = n, group = occ_year, color =  as.factor(occ_year))) +
  geom_line(linewidth = 0.5) +
  theme_minimal() +
  labs(x = "Occurrence month", y = "Number of Occurrence", color = "Occurence year") +
  theme(axis.text.x=element_text(size=6.5)) 
```

Other than seasonal pattern, @tbl-location_count demonstrates the relationship between the location type and number of bicycle theft for each year. It clearly shows that bicycles were mostly stolen when parked outside, and this aligns with common sense that outdoor parking is more vulnerable to theft. However, it is surprising that the second most common location for bicycle theft is apartment instead of house.

@fig-location_count_by_year also illustrates that the number of offence happened outside decreased while thefts happened indoor especially in residence increased in recent years. Factors contributing to the phenomenon could be further explored combining with other data. It is noticeable that there are many reports with location type `Other`. Since it is not specified and the amount of `Other` type location cannot be neglected, how we treat the data may influence the analysis, and there may exists possible reporting bias.

```{r}
#| label: tbl-location_count
#| tbl-cap: Number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
#| echo: false
#| warning: false
#| message: false

data |> group_by(location) |>
  summarise(n=n()) |>
  kable(col.names = c("Location type","Total number of occurrence"))
```
```{r}
#| label: fig-location_count_by_year
#| fig-cap: Yearly number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
#| echo: false
#| warning: false
#| message: false

data |>
  group_by(location, occ_year) |>
  summarise(n=n()) |>
  ggplot(aes(x=location, y=n, group=occ_year, color=as.factor(occ_year))) +
  geom_line() +
  labs(x = "Location Type", y = "Number of occurrence", color = "Occurrence Year") +
  scale_x_discrete() +
  theme_minimal() +
  theme(axis.text.x=element_text(size=6))

```

Other than external pattern, we are also interested how internal variables associated with bicycle thefts. @fig-diff_count_by_year shows the monthly difference in days between occurrence date and report date for each year from 2014 to 2022. It illustrates that people tend to report the offence within the same day of occurrence regardless of years. Very few people report bicycle theft after one month. It also consistent with @fig-yearly_count_by_month that more bicycle theft happen in summer instead of winter. However, the difference in report date and occurrence data does not show a seasonal pattern, namely regardless of seasons, people tend to report within the same day.

```{r}
#| label: fig-diff_count_by_year
#| fig-cap: Monthly difference in days between occurrence date and report date for bicycle thefts in Toronto for each year from 2014 to 2022.
#| echo: false
#| warning: false
#| message: false

data |>
  filter(difference <= 100) |>
  ggplot(aes(x=difference, color=factor(occ_month, levels = month.name))) +
  geom_freqpoly() +
  facet_wrap(~occ_year) +
  theme_minimal() +
  labs(x = "Difference in days", y = "Count", color = "Occurrence month")
```

@tbl-rate_by_year and @tbl-rate_by_month contains recovery rate of bicycle thefts in each year and month respectively. The recovery rate is calculated as the percentage of stolen bicycle that could be successfully returned to the owner. It does not have a clear trend of recovery rate through years, however, 2022 has the lowest recovery rate, which is much less than that in other years (See @tbl-rate_by_year). This is not a good signal that the three lowest recovery rates come from 2022, 2019, 2021, although the highest recovery rate comes from 2020. Such unusual pattern is worth investigating.

@tbl-rate_by_month shows that the recovery rate in winter is lower than that in summer in general. Consider potential seasonal pattern for bicycle thefts in a year, this may also be related to the weather in Toronto that severe and freezing weather in winter constrains people's activity in both positive and negative way. Even recovery rate is higher in summer, it is still not a high value that the largest recovery rate for each year is 17.6% and for each month is 15.7%.

```{r}
#| label: tbl-rate_by_year
#| tbl-cap: Yearly recovery rate of bicycle theft in Toronto from 2014 to 2022.
#| echo: false
#| warning: false
#| message: false

data |> group_by(occ_year) |> 
  summarise(rate = percent(round(sum(bicycle_status == "RECOVERED")/nrow(data)*100,3))) |> 
  arrange(rate) |> 
  kable(col.names = c("Year", "Recovery Rate"))

```

```{r}
#| label: tbl-rate_by_month
#| tbl-cap: Monthly recovery rate of bicycle theft in Toronto from 2014 to 2022
#| echo: false
#| warning: false
#| message: false

data |> group_by(occ_month) |> 
  summarise(rate = (percent(round(sum(bicycle_status == "RECOVERED")/nrow(data)*100,3)))) |> 
  arrange(rate) |> 
  kable(col.names = c("Month", "Recovery Rate"))

```

# Discussion {#sec-discussion}

## Bicycle theft shows an overall decreasing trend in recent years

In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year. Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month. Although the occurrence of bicycle theft decreased recently, bicycle theft is still common in Toronto and it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.

## Bicycle theft shows a seasonal pattern

Possible reasons include improvement of security equipment of the bike, the market behind bicycle theft influenced by pandemic, and people's usage habit of bike. These implications need more data to analyze and could be our further interest.

This may be related to the weather in Toronto that many people will not ride bikes in winter due to extremely freezing weather and frequent snowfall.

## Bicycle theft shows a geographical pattern

Residents' transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips. For residents in house, they may rather choose vehicles that are faster such as cars to commute.

\newpage

# References